Supernovae Time Profiles as a Probe of New Physics at Neutrino Telescopes

### 2024-02-21 paper/sections/conclusion.tex

\textbf{\textit{Results}}---
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/majoran_sensitivity}
    \caption{\textbf{\textit{Exclusion limit for the majoron case.}}
    }
    \label{fig:sensitivity}
\end{figure}
We present the exclusion limit in \fig{sensitivity}.

\textbf{\textit{Outlook}}---

>>>>

\textbf{\textit{Results and Discussions}}---
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/majoran_sensitivity}
    \caption{\textbf{\textit{Exclusion sensitivities for the Majoron case.}}
    The lines on this plot show the exclusion sensitivity for IceCube, and two next-generation neutrino experiments, Hyper-Kamiokande and DUNE.
    Additionally, we also show the reaches by SN cooling (blue) and by the lack of observation of high-energy neutrinos from SN1987A (grey) \cite{Fiorillo:2022cdq}, with the width of the shading region quantifying uncertainties from SN modeling.
    }
    \label{fig:sensitivity}
\end{figure} 
Following our strategies above, we obtain the expected exclusion limit for various BSM if no excess of hits on top of those from background noise and the standard neutrino flux were observed. The expected exclusion limit are presented in \cref{fig:sensitivity} for the Majoron case. To illustrate the broad physics application of the analysis, we also show results for the sterile neutrino case with magnetic portal in the Supplemental Materials. Current limits \cite{Fiorillo:2022cdq} from the energy loss requirement (blue band), and non-observation of high energy neutrinos from SN 1987A (gray band) are also shown in \cref{fig:sensitivity} for the Majoron case. 
With future SN explosion, IceCube can improve existing limits by one orders of magnitude. Moreover, while existing limits from time-integrated observables follow the behavior of $g_\phi\propto m^{-1}_\phi$ as Majoron production rate is proportional to $(g_\phi m_\phi)^2$, IceCube can provide stronger limit than this by focusing on the time window, especially before the peak of SM $\bar{\nu}_e$ fluxes, to reduce SM condemnations.

We also present the estimated reaches of DUNE and Hyper-K in \cref{fig:sensitivity} by looking for time-integrated high energy neutrinos following \cite{Fiorillo:2022cdq, Brdar:2023tmi}. We observe that while for the intermediate mass region, the constraints from IceCube are comparable to that from DUNE and Hyper-K, at both low mass region $m_\phi \lesssim \unit[10]{MeV}$ and high mass region $m_\phi \gtrsim \unit[200]{MeV}$, IceCube can provide stronger constraints. This is the outcome we expect. For light $\phi$ case with negligible time delay, IceCube would observe early signals in a time window separated from that when most of SM flux contribution appears, as we show in the left panel of \cref{fig:hits_and_likelihood}, thus enhancing the reaches. On the high mass end, neutrino signals will arrive much later than the SM case, where a time window at late time can be considered to reduce the standard neutrino contaminations.

Uncertainties on the analysis from different modeling SN explosions are minor, as studied in \cite{li2023old} and potential discrepancies between data and modeling are unresolved questions.
Nevertheless, we point out that the discrepancy could be within $2\sigma$ level which will only change our exclusion limit by a small factor. We also provide the code used in this work so that the impact of SN modeling uncertainties on the sensitivity of neutrino telescopes may be evaluated in the future.

In this work, we have demonstrated that neutrino telescope such as IceCube can provide unprecedented probes of new physics, such as Majoron neutrinos and sterile neutrinos with magnetic portal, via measuring Supernova time profiles to a window as short as $\mathcal{O}(0.01)$ s. Signatures outside the SM neutrino time window, including both early and delayed signals, can be effectively probed at neutrino telescope.
This analysis can also be extended to probing new particles such as sterile neutrinos with mixing portal where the resulting time profile requires more dedicated studies as resonant conversions can happen.

\begin{acknowledgments}
We would like to thank Thomas Janka and Daniel Kresse for providing the data from Garching core collapse SN simulations in machine-readable form, and also for stimulating discussions. Y.-Y. L is supported by the NSF of China through Grant No. 12305107, 12247103.
\end{acknowledgments}

### 2024-02-21 paper/sections/supplements.tex

\section{Appendix A: Dipole Magnetic Moment Portal}
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/magnetic_moment_sensitivity}
    \caption{\textbf{\textit{Exclusion sensitivities for the magnetic moment case.}}
    The lines on this plot show the exclusion sensitivity for IceCube, and two next-generation neutrino experiments, Hyper-Kamiokande and DUNE. The shaded regions are excluded by the energy loss constraints, non-observation of photon and neutrino signals from SN1987A studied in \cite{Brdar:2023tmi} and by   constraints on the energy release from SN explosion following \cite{PhysRevLett.128.221103}.
    }
    \label{fig:magnetic_moment_sensitivity}
\end{figure}
The first scenario under consideration is active-to-sterile neutrino transition magnetic moment described by \cite{Magill:2018jla,Brdar:2020quo,Brdar:2023tmi}
\begin{align}
    \mathcal{L} \supset \sum_\alpha d_\alpha \bar{N}\sigma_{\mu\nu} \nu^{\alpha} F^{\mu\nu}-\frac{M_N}{2} \bar{N}^c N + \text{h.c.}\,,
    \label{eq:Lag}
\end{align}
where $\nu^{\alpha}$ and $N$ represent active and sterile neutrinos, respectively. Further, $F^{\mu\nu}$ is the field strength tensor of the electromagnetic field and $d_\alpha$ is the dimensionful coefficient of this dimension-5 term and $M_N$ is sterile neutrino mass. We will abbreviate $d_\alpha \equiv d$ given that we assume flavor universal interaction. The dominant production channels for $N$ inside the SN are $\nu e^- \to \nu e^-$ at lower energies and $\nu \gamma \to N$ for larger active neutrino energies. Both processes occur due to the interaction term in \cref{eq:Lag}; after $N$ are produced, they decay to active neutrinos and photons which is again realized through the same term in the Lagrangian, with the decay width for $N\to\nu\gamma$ given by $\Gamma_N = 6d^2 M_N^3/4 \pi$ \cite{Plestid:2020vqf}. We notice that the time delay near the exclusion boundary are typically of $\Delta t \sim \mathcal{O}(1)$sec, which overlaps with the standard burst time scale. We don't expect improvements over DUNE and Hyper-K.

{\color{red} Check the bg for DUNE and Hyper-K is the atmospheric neutrino bg which should also contribute to our analysis? IceCube sees at}

>>>>

\section{Appendix A: Dipole Magnetic Moment Portal}
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.47\textwidth]{figures/magnetic_moment_sensitivity}
    \caption{\textbf{\textit{Exclusion sensitivities for the magnetic moment case.}}
    The lines on this plot show the exclusion sensitivity for IceCube, and two next-generation neutrino experiments, Hyper-Kamiokande and DUNE. The shaded regions are excluded by the energy loss constraints, non-observation of photon and neutrino signals from SN1987A studied in \cite{Brdar:2023tmi} and by constraints on the energy release from SN explosion \cite{PhysRevLett.128.221103, Chauhan:2023sci, chauhan2024probing}. \yy{need to update the cyan shaded region as previously excluded using results in \cite{chauhan2024probing}.}
    }
    \label{fig:magnetic_moment_sensitivity}
\end{figure}
The first scenario under consideration is active-to-sterile neutrino transition magnetic moment described by \cite{Magill:2018jla,Brdar:2020quo,Brdar:2023tmi}
\begin{align}
    \mathcal{L} \supset \sum_\alpha d_\alpha \bar{N}\sigma_{\mu\nu} \nu^{\alpha} F^{\mu\nu}-\frac{M_N}{2} \bar{N}^c N + \text{h.c.}\,,
    \label{eq:Lag}
\end{align}
where $\nu^{\alpha}$ and $N$ represent active and sterile neutrinos, respectively. Further, $F^{\mu\nu}$ is the field strength tensor of the electromagnetic field and $d_\alpha$ is the dimensionful coefficient of this dimension-5 term and $M_N$ is sterile neutrino mass. We will abbreviate $d_\alpha \equiv d$ given that we assume flavor universal interaction. We consider the two production channels for $N$ inside the SN: $\nu e^- \to N e^-$ at lower energies and $\nu \gamma \to N$ for larger active neutrino energies. We also point out by neglecting the $N$ production channel $\nu p^+\to N p^+$ inside supernove, the limits we obtained are conservative. Both processes occur due to the interaction term in \cref{eq:Lag}; after $N$ are produced, they decay to active neutrinos and photons which is again realized through the same term in the Lagrangian, with the decay width for $N\to\nu\gamma$ given by $\Gamma_N = 6d^2 M_N^3/4 \pi$ \cite{Plestid:2020vqf}. The limits from IceCube are shown in \cref{fig:magnetic_moment_sensitivity}, together with previously excluded regions by energy loss, non-observations of photons and neutrinos from SN1987A \cite{Brdar:2023tmi}, and constraints on the energy release from SN explosions \cite{PhysRevLett.128.221103,chauhan2024probing}.
We notice that the time delay near the exclusion boundary are typically of $\Delta t \sim \mathcal{O}(1)$ sec, which is longer than the standard $\bar{\nu}_e$ peak luminosity time scale and is not expected to improve over DUNE and Hyper-K performance.
